---
title: "Assessing the quality of the reads"
output: 
   html_document:
      number_section: no
      toc_depth: 2
      toc_float: true
params:
    inputfile:
        value: x
---

## General information

```{r, echo = F}
#source("../R/utils.R")
source("utils.R")
library(pheatmap)
data <- getQualStats(params$inputfile)
color <- colorRampPalette(c("white", "red"))( 50 )
color2 <- colorRampPalette(c("white", "blue"))( 50 )
#a <- c(2,14,22,27,33,37)
qualities <- ""
for (i in data$qual_tags){
   qualities <- paste0(qualities, sprintf("%d (%s), ",i, intToUtf8(i+33) ))
}

qualities <- substr(qualities,1,nchar(qualities)-2) 

```

                    
 Var  | value
----- |------------------------------------
Input file name  | `r sub('.*\\/', '', params$inputfile)`
Read length      | `r data$read_len `
Min good quality | `r data$minQ`
Number of reads  | `r data$nreads`   
Number of highQ reads  | `r data$reads_MlowQ[1]`   
Number of tiles  | `r data$ntiles`
Qualities        | `r qualities`
Reads with N's   | `r data$reads_wN`
Number of N's    | `r sum(data$ACGT_pos[5,])`





## Per base sequence quality


```{r, echo = F}
  QualAbund <- getQualAbundancies(data)
  sampleQ <- mimic(QualAbund)
  my_plot(sampleQ) 

```


## # reads with at least  `m` low Q nucleotides


```{r, echo =F}
barplot(data$reads_MlowQ[2:length(data$reads_MlowQ)], xlab="m" , names=1:data$read_len)
```

## Low Q nucleotide proportion 


```{r, fig.width=4, fig.height=data$ntiles/6, echo = F}
if(all(data$lowQ_ACGT_tile == 0)){
   cat( "\t\t NO LOW QUALITY NUCLEOTIDES FOUND. NOT PLOTTING HEATMAP\n")
}else{
   pheatmap((data$lowQ_ACGT_tile/data$ACGT_tile)[,1:4],cluster_cols=F, cluster_rows=F, 
         fontsize_row=6, fontsize_col=6,  color = color)
} 
```

## Average quality per position per tile

```{r,fig.width=7, fig.height=data$ntiles/6, echo = F }
tiles_meanQ <- array(dim=c(data$ntiles,data$read_len), 
                     dimnames = list(data$tile_tags, 1:data$read_len))
for(i in 1:data$ntiles){
   for(j in 1:data$read_len){
      tiles_meanQ[i,j] <- weighted.mean(data$qual_tags,data$QPosTile_table[j,,i])
   }
}
if (any(is.nan(tiles_meanQ))){
   cat("\t\t THERE WERE TILES WITH MISSING POSITIONS. PLOTTING THEM BLACK.\n")
   color2[length(color2)] <- "#000000"
   from <- min(tiles_meanQ, na.rm=T); to <- max(tiles_meanQ, na.rm=T)
   breaks <- c(seq(from=from,  to=to, by = (to - from)/(length(color2) -1 )), to+0.01)
   tiles_meanQ[which(is.nan(tiles_meanQ))] = to+0.01 
   pheatmap(tiles_meanQ,cluster_cols=F, cluster_rows=F, 
            fontsize_row=6, fontsize_col=6,color=color2, 
            breaks=breaks)
}else{
   pheatmap(tiles_meanQ,cluster_cols=F, cluster_rows=F, 
         fontsize_row=6, fontsize_col=6,color=color2)

}

```

## Low Q nucleotides proportion per position per tile

```{r,fig.width=7, fig.height=data$ntiles/6, echo = F}
tiles_lowQ <- array(dim=c(data$ntiles,data$read_len), 
                     dimnames = list(data$tile_tags, 1:data$read_len))
lowQ_idx <- which(data$qual_tags < data$minQ)
for(i in 1:data$ntiles){
   for(j in 1:data$read_len){
      tiles_lowQ[i,j] <- sum(data$QPosTile_table[j,lowQ_idx,i])/sum(data$QPosTile_table[j,,i])
   }
}
if(sum(is.nan(tiles_lowQ)) >0) {

   cat("\t\t THERE WERE TILES WITH MISSING POSITIONS. REMOVING THESE.\n")
   tiles_lowQ <- tiles_lowQ[rowSums(is.nan(tiles_lowQ)) == 0,]
}
if (all(tiles_lowQ==0)){
   cat( "\t\t NO LOW QUALITY NUCLEOTIDES FOUND. NOT PLOTTING HEATMAP.\n")
}else { 
   pheatmap(tiles_lowQ,cluster_cols=F, cluster_rows=F, 
         fontsize_row=6, fontsize_col=6,color=color)

} 
```


## Low Q nucleotides proportion per position for all tiles
```{r, echo = F, fig.height=2.5 }
N = length(data$qual_tags)

lowQ_prop <- array(dim=c(3,data$read_len), 
                     dimnames = list(c("< 27","< 33","< 37" ), 1:data$read_len))

i = 1

for (lowQ in c(27,33,37)){
   lowQ_idx <- which(data$qual_tags< lowQ)
   for (j in 1:data$read_len){
       lowQ_prop[i,j] <- sum(data$QPosTile_table[j,lowQ_idx,])/sum(data$QPosTile_table[j,,])
   }
   i = i + 1
}

if(all(lowQ==0)){
   cat( "\t\t NO LOW QUALITY NUCLEOTIDES FOUND. NOT PLOTTING HEATMAP.\n")

} else {
   sec_max <- unique(lowQ_prop[3,])[order(unique(lowQ_prop[3,]))[2]]
   breaks <- seq(0,sec_max+0.1,length=length(color))
   breaks <- c(breaks,1)
   pheatmap(lowQ_prop,cluster_cols=F, cluster_rows=F,  
         fontsize_row=6, fontsize_col=6,color=color,breaks = breaks, 
         cellheight=30)
                     
}

```

## Nucleotide content per position

```{r, echo = F}
   x <- 1:data$read_len
   nreads_pos <- colSums(data$ACGT_pos)
   plot(x,data$ACGT_pos[1,]/nreads_pos,type="l",lwd=1.5,col="green",
       xlab="Position in read",ylab="nucleotide content" )
   lines(x, data$ACGT_pos[2,]/nreads_pos,type="l",lwd=1.5,col="blue")
   lines(x, data$ACGT_pos[3,]/nreads_pos,type="l",lwd=1.5,col="black")
   lines(x, data$ACGT_pos[4,]/nreads_pos,type="l",lwd=1.5,col="red")
   lines(x, data$ACGT_pos[5,]/nreads_pos,type="l",lwd=1.5,col="grey")
   legend('topright', c("A","C","G", "T", "N "), bty= "n",lwd=1.5, 
          col=c("green","blue","black","red","grey"))
```

